---
title: "Statistical Analysis Excercise 1"
author: "Rohit Gupta"
date: "June 2, 2017"
output: html_document
---
##Overview
This is part 1 of Coursera Statistical Inference course project. In this simulation exercise, the distribution of averages of 40 exponentials with a thousand simulations is investigated and the exponential distribution in R is compared with the Central Limit Theorem. With this simulation, the following things are shown:

the sample mean and the theoretical mean
the sample variance and the theoretical variance
the distribution
Requirements for this report:

```{r }
library(knitr)
library(ggplot2)
# knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.path='part1/', fig.width=10, fig.height=6, cache=TRUE)
# seed
set.seed(12345)
```

##Simulation
The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations.

```{r }
lambda <- 0.2
sample_size <- 40 
simulations <- 1000
sim_matrix <- matrix(rexp(simulations*sample_size, rate=lambda), simulations, sample_size)
```

The simulation mean can now be plotted below to further explore.
```{r }
sim_mean <- rowMeans(sim_matrix)
hist(sim_mean, xlab="mean of 40 exponentials", ylab="frequence", main="Histogram of the mean of 40 exponentials", col="green")
```

##Sample Mean vs. Theoretical Mean

```{r }
sample_mean <- mean(sim_mean)
cat("sample mean: ", sample_mean)
theoretical_mean <- 1/lambda
cat("theoretical mean: ", theoretical_mean)
```

As shown, the sample mean is very close to the expected theoretical mean.

##Sample Variance vs. Theoretical Variance

```{r }
sample_var <- var(sim_mean)
cat("sample variance: ", sample_var)
```

```{r }
theoretical_var <- (1/lambda)^2/sample_size  
cat("theoretical variance: ", theoretical_var)
```
Again they are very close.

##Distribution
First, we create an approximate normal distribution and see how the sample aligns with it

```{r }
hist(sim_mean, density=20, breaks=20, prob=TRUE, xlab="mean of 40 exponentials", ylab="frequence", main="Histogram of the mean of 40 exponentials", col="grey")
curve(dnorm(x, mean=sample_mean, sd=sqrt(sample_var)), col="black", lwd=2, lty = "dotted", add=TRUE, yaxt="n")
curve(dnorm(x, mean=theoretical_mean, sd=sqrt(theoretical_var)), col="red", lwd=2, add=TRUE, yaxt="n")
```
The plot above shows the histogram can be approximated with the normal distribution.

Second, we have proved that the mean and variance of sample data close to those of normal distribution. Below shows the confidence intervals:

```{r }
sample_conf_interval <- round (mean(sim_mean) + c(-1,1)*1.96*sd(sim_mean)/sqrt(sample_size),3)
cat("sample confidence interval: ", sample_conf_interval)
```

```{r }
theory_conf_interval <- theoretical_mean + c(-1,1)*1.96*sqrt(theoretical_var)/sqrt(sample_size);
cat("theoretical confidence interval: ", theory_conf_interval)
```

Third, plot Q-Q for quantiles

The sample quantiles also match closely with the theorietical quantiles.

```{r }
qqnorm(sim_mean, main="Normal Q-Q Plot", xlab="Theoretical Quantiles", ylab="Sample Quantiles")
qqline(sim_mean, col="blue")
```

